---
title: "Linear regression"
format: html
embed-resources: true
editor: source
---

```{r}
#| label: load-packages
#| include: false

library(here)
library(tidyverse)
library(kableExtra)
library(texreg)
library(plm)
```

## Linear regression

```{r}
finaldata <- read.csv(here("data", "analytical", "finaldata.csv"), header = TRUE)
head(finaldata)
names(finaldata)
```

The model presented on page 6 in the paper is

$$
\text{Outcome}_{ij} = \beta_{0} + \beta_{1}\text{Conflict}_{i(t-1)} + \beta_2\text{Covariates}_{it} + i + t + u_{it}
$$

where $i$ is the country, $t$ is the year, and $u_{it}$ is the error term. We will first use maternal mortality `MatMor` as the $\text{Outcome}_{ij}$ and binary armed conflict specification `armconf1` as the main predictor $\text{Conflict}_{i(t-1)}$. The set of $\text{Covariates}_{it}$ are those included in Table 2. By the way, I think the authors made a careless mistake in writing out the model here. Since there are multiple covariates, $\beta_2$ should be presented as a vector...

```{r}
#| output: asis

finaldata$pctpopdens <- finaldata$popdens / 100
matmormod <- lm(matmor ~ -1 + armconf1 + gdp1000 + OECD + pctpopdens + urban + 
                  agedep + male_edu + temp + rainfall1000 + earthquake + drought + 
                  ISO, data = finaldata)
matmorplm <- plm(matmor ~ armconf1 + gdp1000 + OECD + pctpopdens + urban + 
                  agedep + male_edu + temp + rainfall1000 + earthquake + drought,
                 index = c("ISO"),
                 model = "within",
                 data = finaldata)
htmlreg(list(matmormod, matmorplm))
```

```{r}
#| output: asis

matmormod <- lm(matmor ~ -1 + armconf1 + gdp1000 + OECD + pctpopdens + urban + 
                  agedep + male_edu + temp + rainfall1000 + earthquake + drought + 
                  ISO + as.factor(year), data = finaldata)
matmorplm <- plm(matmor ~ armconf1 + gdp1000 + OECD + pctpopdens + urban + 
                  agedep + male_edu + temp + rainfall1000 + earthquake + drought,
                 index = c("ISO", "year"),
                 effect = "twoways",
                 model = "within",
                 data = finaldata)
htmlreg(list(matmormod, matmorplm))
```

What is the relationship between `matmor` and `gdp1000`?

```{r}
finaldata |>
  dplyr::filter(year < 2018) |>
ggplot(aes(y = matmor, x= gdp1000)) + 
  geom_point() +
  facet_wrap(~ year)
```

For example, if we fit a linear regression line for year 2017.

```{r}
finaldata |>
  dplyr::filter(year == 2017) |>
  ggplot(aes(y = matmor, x= gdp1000)) + 
  geom_point() +
  geom_smooth(method='lm')
```

Take the log of GDP

```{r}
finaldata$loggdp <- log(finaldata$gdp1000)
finaldata |>
  dplyr::filter(year < 2018) |>
ggplot(aes(y = matmor, x= loggdp)) + 
  geom_point() +
  facet_wrap(~ year)
```

```{r}
finaldata$loggdp <- log(finaldata$gdp1000)
matmormod <- lm(matmor ~ -1 + armconf1 + loggdp + OECD + pctpopdens + urban + 
                  agedep + male_edu + temp + rainfall1000 + earthquake + drought + 
                  ISO + as.factor(year), data = finaldata)
summary(matmormod)
matmorplm <- plm(matmor ~ armconf1 + loggdp + OECD + pctpopdens + urban + 
                  agedep + male_edu + temp + rainfall1000 + earthquake + drought,
                 index = c("ISO", "year"),
                 effect = "twoways",
                 model = "within",
                 data = finaldata)
summary(matmorplm)
screenreg(list(matmormod, matmorplm))
```

Now we want to fit the model to the other outcomes: `Un5Mor`, `InfMor`, and `NeoMor`. Since we are using the same predictors, it will be efficient to create an object that lists all the predictors and use the object in `lm()` with the `update.formula()` function to run each regression model with different outcomes.

```{r}
preds <- as.formula(" ~ armconf1 + gdp1000 + OECD + pctpopdens + urban + 
                  agedep + male_edu + temp + rainfall1000 + earthquake + drought + 
                  ISO + as.factor(year)")

matmormod <- lm(update.formula(preds, matmor ~ .), data = finaldata)
un5mormod <- lm(update.formula(preds, un5mor ~ .), data = finaldata)
infmormod <- lm(update.formula(preds, infmor ~ .), data = finaldata)
neomormod <- lm(update.formula(preds, neomor ~ .), data = finaldata)
```

```{r}
preds <- as.formula(" ~ armconf1 + loggdp + OECD + pctpopdens + urban + 
                  agedep + male_edu + temp + rainfall1000 + earthquake + drought")

matmormod <- plm(update.formula(preds, matmor ~ .), index = c("ISO", "year"), effect = "twoways",
                 model = "within", data = finaldata)
un5mormod <- plm(update.formula(preds, un5mor ~ .), index = c("ISO", "year"), effect = "twoways",
                 model = "within", data = finaldata)
infmormod <- plm(update.formula(preds, infmor ~ .), index = c("ISO", "year"), effect = "twoways",
                 model = "within", data = finaldata)
neomormod <- plm(update.formula(preds, neomor ~ .), index = c("ISO", "year"), effect = "twoways",
                 model = "within", data = finaldata)
screenreg(list(matmormod, un5mormod, infmormod, neomormod), ci.force = TRUE,)
```

Then use htmlreg() function from texreg package to create a table of regression coefficients with 95% confidence intervals. But we only want to show select variables so create a list with new variable names.

```{r}
#| output: asis

keepvars <- list("armconf1" = "Armed conflict",
                 "loggdp" = "log(GDP)",
                 "OECD" = "OECD",
                 "pctpopdens" = "Population density",
                 "urban" = "Urban",
                 "agedep" = "Age dependency",
                 "male_edu" = "Male education",
                 "temp" = "Average temperature",
                 "rainfall" = "Average rainfall",
                 "earthquake" = "Earthquake",
                 "drought" = "Drought")
library(texreg)
htmlreg(list(matmormod, un5mormod, infmormod, neomormod), 
        ci.force = TRUE,
        custom.coef.map = keepvars,
        custom.model.names = c("Maternal mortality", "Under-5 mortality",
                               "Infant mortality", "Neonatal mortality"))
```
